---
title: "Missingness"
author: "Guannan Shen"
date: "September 18, 2019"
output: 
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 5
    toc_float: no
  word_document:
    toc: yes
    toc_depth: '5'
  pdf_document:
    latex_engine: lualatex
    number_sections: yes
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE, cache = FALSE}
require("knitr")
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
opts_chunk$set(engine = "R")
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
## setting wd in DELL
## opts_knit$set(root.dir = "~/Stats/CIDA_OMICs/CIDA_OMICS/7659Stats_Genetics/HW5/")
## setting working directory in asus 
## opts_knit$set(root.dir = "C:/Users/hithr/Documents/Stats/gitlab/Cario_RNASeq_Microbiom_Inte/DataRaw/") 
## setting working directory in ubuntu
opts_knit$set(root.dir = "~/Documents/gitlab/ECCHO_github/DataProcessed/more_chems/")
                                                 
## cache = F, if cache = T, will not revaluate code chunk everytime
## double or more space to insert a line break
```


```{r libs, echo = FALSE, warning= F}
######################
## Set up workspace
######################
rm(list = ls())
library(tidyverse)
library(magrittr)
library(BiocStyle)
library(mice)

options(stringsAsFactors = F)
options(dplyr.width = Inf)
getwd()

# directory, Ubuntu 
dir = "~/Documents/gitlab/Omics_Integration/"
# small functions, %nin%, %||%, isnothing
source( paste0(dir, "Code/small_fun.R") )

# ######## clean memory ######################
# rm(list = ls())
# gc()
# is(dds)
# slotNames(dds)
```


# Preprocessing
## Import data and Features
$$N = 407$$ for both of phthalates and phenols.  
Samples should set to missing based on the unadjusted measurements, which are the variables ending in “_ng_ml” (or “nmol_mL”).    
 All other analyses should use the creatinine-adjusted measurements, which are the variables ending in “_ug_g_creat” (or “nmol_g_creat”).  
There are 17 chemicals: 8 phenols and 9 phthalates.   


```{r import}
dir = "~/Documents/gitlab/ECCHO_github/"
## reference plots 
source( paste0(dir, "Code/more_chems/ref_plots.R") )
source( paste0(dir, "Code/more_chems/sum_tables.R") )
source( paste0(dir, "Code/more_chems/Dimen_reduc_clustering.R") )
## import data
phthal_di2_all <- read.csv(paste0(dir,"DataRaw/more_chems/phthal_di2_sep2019.csv"), 
                            header=TRUE)
## keep all chemical data, adjusted and unadjusted 
all_chem <- phthal_di2_all[, -c(2:17)] 
# colnames(all_chem)
dim(all_chem)

## unadjusted for missing 
unadjusted_chem <- all_chem[, 1:19] %>% dplyr::select(-cr_low_flag)
dim(unadjusted_chem)



```

## detection limits and missing 

```{r dect_lim}
colnames(unadjusted_chem)

###### detection limits  #######
dect_lim <- c(0.4, 0.4, 0.3, 0.2, 0.2, 0.2, 
              0.6, 0.2, 0.2, 0.1, 0.1, 1.0,
              0.1, 0.2, 0.1, 0.1, 1.0)

lower_counts <- sapply(colnames(chems), function(x){
  sum(chems[, x] < dect_lim[x])
})
```
