---
title: "Linear Regression CpG & Obesity"
author: "Guannan Shen"
date: "October 2, 2018"
output:
  pdf_document:
    latex_engine: lualatex
    number_sections: yes
    toc: yes
    toc_depth: 5
  word_document:
    toc: yes
    toc_depth: '5'
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE, cache = FALSE}
require("knitr")
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
opts_chunk$set(engine = "R")
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
## setting working directory in asus 
## opts_knit$set(root.dir = "C:/Users/hithr/Documents/Stats/gitlab/ECCHO_Guannan/Code/")
## setting working directory in ubuntu
opts_knit$set(root.dir = "~/Documents/gitlab/ECCHO_github/Code/")
                                                 
## cache = F, if cache = T, will not revaluate code chunk everytime
## double or more space to insert a line break
```


```{r libs}
## set up workspace
library(knitr)
library(tidyverse)
options(stringsAsFactors = F)
options(dplyr.width = Inf)
getwd()
```
# ECCHO_Guannan

ECCHO grant (chemicals and methylation project of the Healthy Start cohort).

## 9/27/2018
linear regression model for 9 outcomes on obesity and CpG  
$ y = CpG + maternal age + race + CellTypes(7) (All Male)$  
The linear models, each has 10 covariates in the   
Outcomes: "birth_weight", "ipv3_pp_fm_pct", "Chol_IPV3", "FFA_IPV3", "Gluc_IPV3", "HDL_IPV3", "Insu_IPV3", "Trig_IPV3", "Leptin_actual__ng_ml_".  
Although we discussed include the variable “infant sex” as a covariate, we actually did the chemicals-methylation analysis stratified by sex. So the DMRs were only detected among male offspring. Therefore we should probably restrict our analysis to individuals with infant_sex = 2 (males), at least to begin with.  
There will be missing data for some outcomes, I think it is OK if the sample size differs for each model.  


## Methods
### Data Pre-processing
The dataset to model the association between obesity and CpG M-value was generated from three dataset: cordblood cell counts (celltype), obesity related outcomes (pfas) and top300 CpG list (m300). The final dataset was the inner join product (the intersection) of these 3 datasets and was joint by "pid" (participants ID). The "celltype" and "m300" datasets share the same collection of participants with sample size 600. The outcome data set contains 589 subjects, and 588 out of 589 subjects are shared across all 3 dataset. The subject with pid 30568 is only in the "pfas" dataset.  
In this way, the final dataset obs_cpg contains 588 rows (subjects) and 320 columns. After the dataset was filtered by "infant_sex == 2 (males)", 308 participants left.  
Regarding the male only dataset, "cpg_male", there are 308 rows and 320 columns. Missing values only occur in outcomes, thus there are different sample sizes for different outcomes. The first 2 columns, pid and infant_sex would not be included in the linear regression model. The 3rd and 4th columns are "maternal_age" and categorical variable "race_4", which would always be included in the model. From column 5 to 13, these are 9 outcomes. From column 14 to 20, these are 7 celltypes, "Bcell", "CD4T", "CD8T", "Gran", "Mono", "NK" and "nRBC",  which would always be included in the model. From column 21 to 320, these are 300 CpG M values. Each linear regression model would contain 1 out of 300 CpGs. 

```{r data process}
# readin data in ASUS
# largest dataset is 3Mb
# celltype <- read_csv("C:/Users/hithr/Documents/Stats/gitlab/ECCHO_Guannan/DataProcessed/healthy_start_cordblood_cellcounts_id.txt")
# m300 <- read_csv("C:/Users/hithr/Documents/Stats/gitlab/ECCHO_Guannan/DataRaw/HS_450K_CB_Mval_normbatch_StarlingSubset_10-01-18.csv")
# pfas <- read_csv("C:/Users/hithr/Documents/Stats/gitlab/ECCHO_Guannan/DataRaw/pfas_methyl_di.csv")

## readin data in Ubuntu
celltype <- read_csv("~/Documents/gitlab/ECCHO_github/DataProcessed/healthy_start_cordblood_cellcounts_id.txt")

m300 <- read_csv("~/Documents/gitlab/ECCHO_github/DataRaw/HS_450K_CB_Mval_normbatch_StarlingSubset_10-01-18.csv")
pfas <- read_csv("~/Documents/gitlab/ECCHO_github/DataRaw/pfas_methyl_di.csv")

# Exploratory Data Analysis
dim(celltype)
colnames(celltype)
##
dim(m300)
##
summary(pfas)
dim(pfas)
outcomes <- colnames(pfas)[-c(1,3:5)]
paste(outcomes)

# compare m300 and celltype, the subjects
m300$pid[!(m300$pid == celltype$pid)]
sum(m300$pid %in% celltype$pid)


## merge dataset by ID, inner join
pfas_cell <- merge(pfas, celltype, by = "pid")
obs_cpg <- merge(pfas_cell, m300, by = "pid")
dim(obs_cpg)

# check final sujects with outcomes
pfas$pid[!(obs_cpg$pid == pfas$pid)]

#
pfas$pid[!(pfas$pid %in% m300$pid)]

message("The subject with PID 30568 is in Outcomes but not in celltype and top 300 CpG")


# filter based on gender, make race_4 as catrgorical variable. 
cpg_male <- obs_cpg %>% filter(infant_sex == 2) %>% 
  mutate(race_4 = as.factor(race_4) ) %>% 
  select(pid, infant_sex, maternal_age, race_4, everything())

dim(cpg_male)

############### missing data  ####################
kable(apply(cpg_male[, 5:13], 2, function(x){sum(is.na(x))}) , 
      caption = "Missing Data Summary",
      col.names = "# Missing Values")


##################################################
## to build up the lm model, get the variables' name

## "maternal_age", "race_4" 
colnames(cpg_male)[3:4]

## outcomes
## "birth_weight", "ipv3_pp_fm_pct", "Chol_IPV3", "FFA_IPV3", "Gluc_IPV3", "HDL_IPV3", "Insu_IPV3"             "Trig_IPV3", "Leptin_actual__ng_ml_"
colnames(cpg_male)[5:13]

## cell types names
## "Bcell", "CD4T", "CD8T", "Gran", "Mono", "NK" and "nRBC" 
colnames(cpg_male)[14: 20]

message("loop over 21 to 320")



```


```{r looplm, echo = FALSE}
################# loop for birth_weight #################################
birth_weight_lm <- lapply(21:320, function(i){
  lm = lm(birth_weight ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
birth_weight_lm <- data.frame(matrix(unlist(birth_weight_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
birth_weight_lm <-  birth_weight_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything())
birth_weight_lm <- birth_weight_lm[order(birth_weight_lm$FDR), ]

## sample size 
size1 <- length(cpg_male$birth_weight) - sum(is.na(cpg_male$birth_weight))

## summary table 
kable(head(birth_weight_lm,10), caption = paste("Top10 FDR birth_weight ", "n = ", size1, sep = ""))




####################### ipv3_pp_fm_pct ####################
ipv3_pp_fm_pct_lm <- lapply(21:320, function(i){
  lm = lm(ipv3_pp_fm_pct ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
ipv3_pp_fm_pct_lm <- data.frame(matrix(unlist(ipv3_pp_fm_pct_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
ipv3_pp_fm_pct_lm <-  ipv3_pp_fm_pct_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 

ipv3_pp_fm_pct_lm <- ipv3_pp_fm_pct_lm[order(ipv3_pp_fm_pct_lm$FDR), ]

## sample size 
size2 <- length(cpg_male$ipv3_pp_fm_pct) - sum(is.na(cpg_male$ipv3_pp_fm_pct))

## summary table 
kable(head(ipv3_pp_fm_pct_lm,10), caption = paste("Top10 FDR ipv3_pp_fm_pct ", "n = ", size2, sep = ""))

########################## Chol_IPV3 ###############################
Chol_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(Chol_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
Chol_IPV3_lm <- data.frame(matrix(unlist(Chol_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
Chol_IPV3_lm <-  Chol_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 
Chol_IPV3_lm <- Chol_IPV3_lm[order(Chol_IPV3_lm$FDR), ]


## sample size 
size3 <- length(cpg_male$Chol_IPV3) - sum(is.na(cpg_male$Chol_IPV3))

## summary table 
kable(head(Chol_IPV3_lm,10), caption = paste("Top10 FDR Chol_IPV3 ", "n = ", size3, sep = ""))

###################### FFA_IPV3 ##################################
FFA_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(FFA_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
FFA_IPV3_lm <- data.frame(matrix(unlist(FFA_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
FFA_IPV3_lm <-  FFA_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 
FFA_IPV3_lm <- FFA_IPV3_lm[order(FFA_IPV3_lm$FDR), ]


## sample size 
size4 <- length(cpg_male$FFA_IPV3) - sum(is.na(cpg_male$FFA_IPV3))

## summary table 
kable(head(FFA_IPV3_lm, 10), caption = paste("Top10 FDR FFA_IPV3 ", "n = ", size4, sep = ""))

####################### Gluc_IPV3 ##########################
Gluc_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(Gluc_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
Gluc_IPV3_lm <- data.frame(matrix(unlist(Gluc_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
Gluc_IPV3_lm <-  Gluc_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 
Gluc_IPV3_lm <- Gluc_IPV3_lm[order(Gluc_IPV3_lm$FDR), ]

## sample size 
size5 <- length(cpg_male$Gluc_IPV3) - sum(is.na(cpg_male$Gluc_IPV3))

## summary table 
kable(head(Gluc_IPV3_lm, 10), caption = paste("Top10 FDR Gluc_IPV3 ", "n = ", size5, sep = ""))


##################### HDL_IPV3 ######################
HDL_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(HDL_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
HDL_IPV3_lm <- data.frame(matrix(unlist(HDL_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
HDL_IPV3_lm <-  HDL_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything())
HDL_IPV3_lm <- HDL_IPV3_lm[order(HDL_IPV3_lm$FDR), ]


## sample size 
size6 <- length(cpg_male$HDL_IPV3) - sum(is.na(cpg_male$HDL_IPV3))

## summary table 
kable(head(HDL_IPV3_lm, 10), caption = paste("Top10 FDR HDL_IPV3 ", "n = ", size6, sep = ""))

################### Insu_IPV3 ###############################
Insu_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(Insu_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
Insu_IPV3_lm <- data.frame(matrix(unlist(Insu_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
Insu_IPV3_lm <-  Insu_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 
Insu_IPV3_lm <- Insu_IPV3_lm[order(Insu_IPV3_lm$FDR), ]

## sample size 
size7 <- length(cpg_male$Insu_IPV3) - sum(is.na(cpg_male$Insu_IPV3))

## summary table 
kable(head(Insu_IPV3_lm, 10), caption = paste("Top10 FDR Insu_IPV3 ", "n = ", size7, sep = ""))

############## Trig_IPV3 #########################
Trig_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(Trig_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
Trig_IPV3_lm <- data.frame(matrix(unlist(Trig_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
Trig_IPV3_lm <-  Trig_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 
Trig_IPV3_lm <- Trig_IPV3_lm[order(Trig_IPV3_lm$FDR), ]


## sample size 
size8 <- length(cpg_male$Trig_IPV3) - sum(is.na(cpg_male$Trig_IPV3))

## summary table 
kable(head(Trig_IPV3_lm, 10), caption = paste("Top10 FDR Trig_IPV3 ", "n = ", size8, sep = ""))

############# Leptin_actual__ng_ml_ #####################
Leptin_actual__ng_ml__lm <- lapply(21:320, function(i){
  lm = lm(Leptin_actual__ng_ml_ ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
Leptin_actual__ng_ml__lm <- data.frame(matrix(unlist(Leptin_actual__ng_ml__lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
Leptin_actual__ng_ml__lm <-  Leptin_actual__ng_ml__lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything())

Leptin_actual__ng_ml__lm <- Leptin_actual__ng_ml__lm[order(Leptin_actual__ng_ml__lm$FDR), ]


## sample size 
size9 <- length(cpg_male$Leptin_actual__ng_ml_) - sum(is.na(cpg_male$Leptin_actual__ng_ml_))

## summary table 
kable(head(Leptin_actual__ng_ml__lm, 10), caption = paste("Top10 FDR Leptin_actual__ng_ml ", "n = ", size9, sep = ""))

### Double check 
check_lm <- lm(Leptin_actual__ng_ml_ ~ cpg_male[ , 42] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
summary(check_lm)


```

```{r code, eval = FALSE}
################# loop for birth_weight #################################
birth_weight_lm <- lapply(21:320, function(i){
  lm = lm(birth_weight ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
birth_weight_lm <- data.frame(matrix(unlist(birth_weight_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
birth_weight_lm <-  birth_weight_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything())
birth_weight_lm <- birth_weight_lm[order(birth_weight_lm$FDR), ]

## sample size 
size1 <- length(cpg_male$birth_weight) - sum(is.na(cpg_male$birth_weight))

## summary table 
kable(head(birth_weight_lm,10), caption = paste("Top10 FDR birth_weight ", "n = ", size1, sep = ""))




####################### ipv3_pp_fm_pct ####################
ipv3_pp_fm_pct_lm <- lapply(21:320, function(i){
  lm = lm(ipv3_pp_fm_pct ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
ipv3_pp_fm_pct_lm <- data.frame(matrix(unlist(ipv3_pp_fm_pct_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
ipv3_pp_fm_pct_lm <-  ipv3_pp_fm_pct_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 

ipv3_pp_fm_pct_lm <- ipv3_pp_fm_pct_lm[order(ipv3_pp_fm_pct_lm$FDR), ]

## sample size 
size2 <- length(cpg_male$ipv3_pp_fm_pct) - sum(is.na(cpg_male$ipv3_pp_fm_pct))

## summary table 
kable(head(ipv3_pp_fm_pct_lm,10), caption = paste("Top10 FDR ipv3_pp_fm_pct ", "n = ", size2, sep = ""))

########################## Chol_IPV3 ###############################
Chol_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(Chol_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
Chol_IPV3_lm <- data.frame(matrix(unlist(Chol_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
Chol_IPV3_lm <-  Chol_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 
Chol_IPV3_lm <- Chol_IPV3_lm[order(Chol_IPV3_lm$FDR), ]


## sample size 
size3 <- length(cpg_male$Chol_IPV3) - sum(is.na(cpg_male$Chol_IPV3))

## summary table 
kable(head(Chol_IPV3_lm,10), caption = paste("Top10 FDR Chol_IPV3 ", "n = ", size3, sep = ""))

###################### FFA_IPV3 ##################################
FFA_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(FFA_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
FFA_IPV3_lm <- data.frame(matrix(unlist(FFA_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
FFA_IPV3_lm <-  FFA_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 
FFA_IPV3_lm <- FFA_IPV3_lm[order(FFA_IPV3_lm$FDR), ]


## sample size 
size4 <- length(cpg_male$FFA_IPV3) - sum(is.na(cpg_male$FFA_IPV3))

## summary table 
kable(head(FFA_IPV3_lm, 10), caption = paste("Top10 FDR FFA_IPV3 ", "n = ", size4, sep = ""))

####################### Gluc_IPV3 ##########################
Gluc_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(Gluc_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
Gluc_IPV3_lm <- data.frame(matrix(unlist(Gluc_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
Gluc_IPV3_lm <-  Gluc_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 
Gluc_IPV3_lm <- Gluc_IPV3_lm[order(Gluc_IPV3_lm$FDR), ]

## sample size 
size5 <- length(cpg_male$Gluc_IPV3) - sum(is.na(cpg_male$Gluc_IPV3))

## summary table 
kable(head(Gluc_IPV3_lm, 10), caption = paste("Top10 FDR Gluc_IPV3 ", "n = ", size5, sep = ""))


##################### HDL_IPV3 ######################
HDL_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(HDL_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
HDL_IPV3_lm <- data.frame(matrix(unlist(HDL_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
HDL_IPV3_lm <-  HDL_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything())
HDL_IPV3_lm <- HDL_IPV3_lm[order(HDL_IPV3_lm$FDR), ]


## sample size 
size6 <- length(cpg_male$HDL_IPV3) - sum(is.na(cpg_male$HDL_IPV3))

## summary table 
kable(head(HDL_IPV3_lm, 10), caption = paste("Top10 FDR HDL_IPV3 ", "n = ", size6, sep = ""))

################### Insu_IPV3 ###############################
Insu_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(Insu_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
Insu_IPV3_lm <- data.frame(matrix(unlist(Insu_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
Insu_IPV3_lm <-  Insu_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 
Insu_IPV3_lm <- Insu_IPV3_lm[order(Insu_IPV3_lm$FDR), ]

## sample size 
size7 <- length(cpg_male$Insu_IPV3) - sum(is.na(cpg_male$Insu_IPV3))

## summary table 
kable(head(Insu_IPV3_lm, 10), caption = paste("Top10 FDR Insu_IPV3 ", "n = ", size7, sep = ""))

############## Trig_IPV3 #########################
Trig_IPV3_lm <- lapply(21:320, function(i){
  lm = lm(Trig_IPV3 ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
Trig_IPV3_lm <- data.frame(matrix(unlist(Trig_IPV3_lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
Trig_IPV3_lm <-  Trig_IPV3_lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything()) 
Trig_IPV3_lm <- Trig_IPV3_lm[order(Trig_IPV3_lm$FDR), ]


## sample size 
size8 <- length(cpg_male$Trig_IPV3) - sum(is.na(cpg_male$Trig_IPV3))

## summary table 
kable(head(Trig_IPV3_lm, 10), caption = paste("Top10 FDR Trig_IPV3 ", "n = ", size8, sep = ""))

############# Leptin_actual__ng_ml_ #####################
Leptin_actual__ng_ml__lm <- lapply(21:320, function(i){
  lm = lm(Leptin_actual__ng_ml_ ~ cpg_male[ ,i] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
  coef = round(summary(lm)$coefficients[2, ],4)
  return(coef)
})
Leptin_actual__ng_ml__lm <- data.frame(matrix(unlist(Leptin_actual__ng_ml__lm), ncol = 4, byrow = TRUE,
                          dimnames = list(
                     c(colnames(cpg_male)[21:320]),
                        c("Estimate", "Std.Error", "t.statistic", "p.value"))))

# adjusted p-value
Leptin_actual__ng_ml__lm <-  Leptin_actual__ng_ml__lm %>% mutate(FDR = p.adjust(p.value, "BH", 300 ),
                                               FWER = p.adjust(p.value,"bonferroni", 300),
                                               names = colnames(cpg_male)[21:320] ) %>% 
                                               select(names, everything())

Leptin_actual__ng_ml__lm <- Leptin_actual__ng_ml__lm[order(Leptin_actual__ng_ml__lm$FDR), ]


## sample size 
size9 <- length(cpg_male$Leptin_actual__ng_ml_) - sum(is.na(cpg_male$Leptin_actual__ng_ml_))

## summary table 
kable(head(Leptin_actual__ng_ml__lm, 10), caption = paste("Top10 FDR Leptin_actual__ng_ml ", "n = ", size9, sep = ""))

### Double check 
check_lm <- lm(Leptin_actual__ng_ml_ ~ cpg_male[ , 42] + maternal_age + race_4 + Bcell + CD4T +
            CD8T + Gran + Mono + NK + nRBC , data = cpg_male)
summary(check_lm)



```
  
  
  
  