---
title: "StarlingExtractCpGNames"
author: "Harry Smith"
date: "10/1/2018"
output: html_document
---
## Introduction

The purpose of this program is extract a list of the most significant CpGs for
each significant (FDR < 0.05) DMR. We will only select CpGs from male subjects. 
The raw data was sent to Harry Smith by Weiming and received on 10/01/18. Harry
extracted the list of most significant CpG names based on the p - value for 
males which can be found in the column labeled 'P value M'. This program will also 
contain the code used to subset the matrix of M-values for the Cord blood in the
Healthy Start Cohort. This code was run on the Lead server. 

Data in and out for CpG list generation - Local machine
Input data: "/Users/harry/Documents/CIDA/Projects/Starling/dataRaw/CpG_from_DMR_adjusted_M.txt"
Output data: "/Users/harry/Documents/CIDA/Projects/Starling/dataProcessed/CpGList_Males.txt"

```{r}
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
'%!in%' <- function(x,y)!('%in%'(x,y))
```

```{r setup}
## set up workspace
library(knitr)
library(dplyr)
library(data.table)
options(stringsAsFactors = F)
options(dplyr.width = Inf)

dir <- "/Users/harry/Documents/CIDA/Projects/Starling/"
draw <- "dataRaw/"
dpro <- "dataProcessed/"
rawdata <- "CpG_from_DMR_adjusted_M.txt"
processed <- "CpGList_Males.txt"
```

```{r data_read}
## read in raw data
cpgRaw <- fread(file = paste0(dir, draw, rawdata))
sum(!duplicated(cpgRaw$dmr_num)) # 300 DMRs

## extract the CpG with the most significant male pvalue - Should have 300 rows
cpgFilt <- cpgRaw %>% 
  group_by(dmr_num) %>%
  slice(which.min(`P value M`))

## put CpG names in a vector
cpgNames <- cpgFilt$CpG

## Save as text file for subsetting m-values on lead server
write.table(cpgNames, file = paste0(dir, dpro, processed), 
            row.names = F, 
            col.names = F, 
            quote = F)
```

Data in and out for M-value subsetting - Lead server
Input data: "/home/datasets/For_DWH/HS_450K_CB_Mval_normbatch_02-10-16.csv"
            "/home/smiharry/ECHO/methylation/data/CpGList_Males.txt"
Output data: "/home/datasets/For_DWH/HS_450K_CB_Mval_normbatch_StarlingSubset_10-01-18.csv"

```{r data_subset, eval = F}
## set up workspace on lead server
library(dplyr)
library(data.table)
options(stringsAsFactors = F)
options(dplyr.width = Inf)

dir <- "/home/datasets/For_DWH/"
cpglist <- "/home/smiharry/ECHO/methylation/data/CpGList_Males.txt"
rawdata <- "HS_450K_CB_Mval_normbatch_02-10-16.csv"
processed <- "HS_450K_CB_Mval_normbatch_StarlingSubset_10-01-18.csv"
## read in M-val matrix
mval <- fread(paste0(dir, rawdata), header = T)
cpg <- read.table(file = cpglist, header = F)

## add pid to cpg names list
cpg <- c("pid", as.character(cpg$V1))

## Filter mvalues based on cpg names from list generated above
mvalfilt <- mval %>%
  select(one_of(cpg))

## Write out dataset
write.csv(mvalfilt, file = paste0("/home/smiharry/ECHO/methylation/data/", processed), row.names = F, quote = T)

```

## Final M-value subset

Per the discussion from our meeting on 10/25/18, we will perform the untargeted
analysis on the CpGs with FDR < 0.05 as reported in the dataset sent from 
Weiming's DMR analysis for both Males and Females. There are 103 CpGs in the 
Male dataset that meet this criteria, and 38 CpGs in the female dataset. Harry
will use the CpG probe IDs from the data sets below to subset the M-value dataset 
on the LEAD server, and transfer this data set to Guannan for the linear regression
analysis.

```{r Final_Data_Subset, eval = FALSE}
## set up workspace on lead server
library(dplyr)
library(data.table)
options(stringsAsFactors = F)
options(dplyr.width = Inf)

dir <- "/home/datasets/For_DWH/"
cpgdir <- "/home/smiharry/ECHO/methylation/data/"
cpgfemale <- "top_CpG_from_DMR_adjusted_F005.txt"
cpgmale <- "top_CpG_from_DMR_adjusted_M005.txt"
rawdata <- "HS_450K_CB_Mval_normbatch_02-10-16.csv"
processed_F <- "HS_450K_CB_Mval_normbatch_StarlingSubset_FEMALE_10-29-18.csv"
processed_M <- "HS_450K_CB_Mval_normbatch_StarlingSubset_MALE_10-29-18.csv"

## read in M-val matrix
mval <- fread(paste0(dir, rawdata), header = T)
## read in CpG data for males and females
cpg_F <- read.delim(file = paste0(cpgdir, cpgfemale), header = T, sep = "\t")
dim(cpg_F)
cpg_M <- read.delim(file = paste0(cpgdir, cpgmale), header = T, sep = "\t")
dim(cpg_M)


## Filter mvalues based on cpg names from list generated above
#Female
mvalfilt_F <- mval %>%
  select(one_of("pid", cpg_F$CpG))
#Male
mvalfilt_M <- mval %>%
  select(one_of("pid", cpg_M$CpG))

## Write out dataset
write.csv(mvalfilt_F, file = paste0("/home/smiharry/ECHO/methylation/data/", processed_F), row.names = F, quote = T)
write.csv(mvalfilt_M, file = paste0("/home/smiharry/ECHO/methylation/data/", processed_M), row.names = F, quote = T)
```


```

